### 2015-05-25 paper/paper.tex
Finished off content of implementation section

\lstdefinelanguage{JavaScript}{
  keywords={typeof, new, true, false, catch, function, delete, return, null, catch, switch, var, if, in, while, do, else, case, break},
  keywordstyle=\color{blue}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this},
  ndkeywordstyle=\color{darkgray}\bfseries,
  identifierstyle=\color{black},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{purple}\ttfamily,
  stringstyle=\color{red}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

\lstset{
   language=JavaScript,
   backgroundcolor=\color{lightgray},
   extendedchars=true,
   basicstyle=\footnotesize\ttfamily,
   showstringspaces=false,
   showspaces=false,
   numbers=left,
   numberstyle=\footnotesize,
   numbersep=9pt,
   tabsize=2,
   breaklines=true,
   showtabs=false,
   captionpos=b
}

@@

The \texttt{-d} and \texttt{-c} flags enable the specifying of database and collection within that database in which to
import the data into. The \texttt{--headerline} flag asks Mongo to use the CSV header to automatically generate BSON
keys for each document. So each row in the CSV file will be converted to a MongoDB document, and stored in the specified
database collection. By default, MongoDB tries to guess the data type for each value, with numbers represented as Doubles.
Of course, storing every document as a flat container for different floating point values is not optimal, and pre-processing
can be performed on Mongo to better structure the documents for later querying and processing.

@@

\begin{lstlisting}
var query = {Track: {$exists: true}};

db.train.find(query).forEach(function(doc) {

  // Cast Track to be Int32 type
  doc.Track = new NumberInt(doc.Track);

  // Write changes to MongoDB
  db.train.save(doc);

});
\end{lstlisting}

@@

\begin{lstlisting}
// Change LoadEmpty = 1 to 'loaded'
db.train.update(

@@

\begin{lstlisting}
// Create an ascending index on LoadEmpty field
db.train.createIndex({LoadEmpty: 1});
\end{lstlisting}

@@

\begin{lstlisting}
{
  "_id" : ObjectId("5561e94530f853aef8bbfd0c"),
  "Time" : 1094468487.0999999046325684,
  "Speed" : 4.4783997535999998,
  "LoadEmpty" : -1.0000000000000000,
  "km" : 5.0096998214999999,
  "Lat" : -20.6665750019999983,
  "Lon" : 117.1280816499999986,
  "Track" : 2.0000000000000000,
  "SND1" : 0.0118295024810000,
  "SND2" : -0.0173957574360000,
  "SND3" : -0.0295975576070000,
  "SND4" : -0.0327308348410000,
  "CouplerForce" : 23.1114991600000010,
  "LateralAccel" : 0.0033679292537000,
  "AccLeft" : 0.0000000000000000,
  "AccRight" : 0.0000000000000000,
  "BounceFront" : -0.0104506661800000,
  "BounceRear" : -0.0234966575210000,
  "RockFront" : 0.0445603373210000,
  "RockRear" : 0.0122018001710000
}
\end{lstlisting}

@@

\begin{lstlisting}
{
  "_id" : ObjectId("555d30ca33d9d7607772920f"),
  "Time" : 1094468487.0999999046325684,
  "Speed" : 4.4783997535999998,
  "LoadEmpty" : "unknown",
  "km" : 5.0096998214999999,
  "Track" : 2,
  "CouplerForce" : 23.1114991600000010,
  "LateralAccel" : 0.0033679292537000,
  "SND" : {
    "SND1" : 0.0118295024810000,
    "SND2" : -0.0173957574360000,
    "SND3" : -0.0295975576070000,
    "SND4" : -0.0327308348410000
  },
  "Loc" : {
    "Lat" : -20.6665750019999983,
    "Lon" : 117.1280816499999986
  },
  "Bounce" : {
    "Front" : -0.0104506661800000,
    "Rear" : -0.0234966575210000
  },
  "Rock" : {
    "Front" : 0.0445603373210000,
    "Rear" : 0.0122018001710000
  },
  "Acc" : {
    "Left" : 0.0000000000000000,
    "Right" : 0.0000000000000000
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
var updateDocsFunc = function(doc) {
  // Add location nested doc
  doc.Loc = {
    Lat: doc.Lat,
    Lon: doc.Lon
  };

@@

When operations are performed on multiple documents in Mongo, if the documents contain a different amount of keys, Mongo
effectively ignores this and continues to perform the operation on all documents. Generally, specific keys and values
are passed to operations such as these, hence if the key used does not exist within certain documents, those documents
simply get ignored from the operation. For example, if a collection contains 100 documents, where only four of those
documents contain the \textbf{isAdmin} key and an operation is called that depends on the key \textbf{isAdmin}, only
those four documents within the collection will be operated on. This allows completely differently structured documents
to be contained within the same collection (however, in a well-structured dataset, multiple collections would be used).

@@

@@

@@

>>>>

\lstdefinelanguage{JavaScript}{
  keywords={typeof, new, true, false, catch, function, delete, return, undefined, null, catch, switch, var, if, in, while, do, else, case, break},
  keywordstyle=\color{blue}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this},
  ndkeywordstyle=\color{darkgray}\bfseries,
  identifierstyle=\color{black},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{purple}\ttfamily,
  stringstyle=\color{red}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

\lstset{
  language=JavaScript,
  backgroundcolor=\color{lightgray},
  extendedchars=true,
  basicstyle=\footnotesize\ttfamily,
  showstringspaces=false,
  showspaces=false,
  numbers=none,
  tabsize=2,
  breaklines=true,
  showtabs=false,
  captionpos=b
}

@@

The \texttt{-d} and \texttt{-c} flags enable the specifying of database and collection within that database in which to
import the data into. The \texttt{--headerline} flag asks Mongo to use the CSV header to automatically generate BSON
keys for each document. The \texttt{--ignoreBlanks} flag asks Mongo not to set empty CSV values to the empty string. Instead,
these key-value pairs are left out of their respective documents.
So each row in the CSV file will be converted to a MongoDB document, and stored in the specified
database collection. By default, MongoDB tries to guess the data type for each value, with numbers represented as Doubles.
Of course, storing every document as a flat container for different floating point values is not optimal, and pre-processing
can be performed on Mongo to better structure the documents for later querying and processing.

@@

\begin{lstlisting}[caption=Performing an iterative operation on database collection ``train''.]
var query = {Track: {$exists: true}};

db.train.find(query).forEach(function(doc) {
  // Cast Track to be Int32 type
  doc.Track = new NumberInt(doc.Track);

  // Write changes to MongoDB
  db.train.save(doc);
});
\end{lstlisting}

@@

\begin{lstlisting}[caption=Performing asynchronous \texttt{update()} operations on ``train'' collection.]
// Change LoadEmpty = 1 to 'loaded'
db.train.update(

@@

\begin{lstlisting}[caption=Creating an index on ``LoadEmpty'' key in ``train'' collection.]
// Create an ascending index on LoadEmpty field
db.train.createIndex({LoadEmpty: 1});
\end{lstlisting}

@@

\begin{lstlisting}[caption=Original ``flat'' document structure.]
{
  "_id" : ObjectId("5561e94530f853aef8bbfd0c"),
  "Time" : 1094468487.0999999046325684,
  "Speed" : 4.4783997535999998,
  "LoadEmpty" : -1.0000000000000000,
  "km" : 5.0096998214999999,
  "Lat" : -20.6665750019999983,
  "Lon" : 117.1280816499999986,
  "Track" : 2.0000000000000000,
  "SND1" : 0.0118295024810000,
  "SND2" : -0.0173957574360000,
  "SND3" : -0.0295975576070000,
  "SND4" : -0.0327308348410000,
  "CouplerForce" : 23.1114991600000010,
  "LateralAccel" : 0.0033679292537000,
  "AccLeft" : 0.0000000000000000,
  "AccRight" : 0.0000000000000000,
  "BounceFront" : -0.0104506661800000,
  "BounceRear" : -0.0234966575210000,
  "RockFront" : 0.0445603373210000,
  "RockRear" : 0.0122018001710000
}
\end{lstlisting}

@@

\begin{lstlisting}[caption=Improved multi-layered document structure.]
{
  "_id" : ObjectId("555d30ca33d9d7607772920f"),
  "Time" : 1094468487.0999999046325684,
  "Speed" : 4.4783997535999998,
  "LoadEmpty" : "unknown",
  "km" : 5.0096998214999999,
  "Track" : 2,
  "CouplerForce" : 23.1114991600000010,
  "LateralAccel" : 0.0033679292537000,
  "SND" : {
    "SND1" : 0.0118295024810000,
    "SND2" : -0.0173957574360000,
    "SND3" : -0.0295975576070000,
    "SND4" : -0.0327308348410000
  },
  "Loc" : {
    "Lat" : -20.6665750019999983,
    "Lon" : 117.1280816499999986
  },
  "Bounce" : {
    "Front" : -0.0104506661800000,
    "Rear" : -0.0234966575210000
  },
  "Rock" : {
    "Front" : 0.0445603373210000,
    "Rear" : 0.0122018001710000
  },
  "Acc" : {
    "Left" : 0.0000000000000000,
    "Right" : 0.0000000000000000
  }
}
\end{lstlisting}

@@

\begin{lstlisting}[caption=Restructuring of documents in ``train'' collection using iterative operation.]
var updateDocsFunc = function(doc) {
  // Add location nested doc
  doc.Loc = {
    Lat: doc.Lat,
    Lon: doc.Lon
  };

@@

When operations are performed on multiple documents in Mongo, if the documents contain a different amount of keys, Mongo
effectively ignores this and continues to perform the operation on all documents. Generally, specific keys and values
are passed to operations such as these, hence if the key used does not exist within certain documents, those documents
simply get ignored from the operation. For example, if a collection contains 100 documents, where only four of those
documents contain the \textbf{isAdmin} key, and an operation is called that depends on the key \textbf{isAdmin}, only
those four documents within the collection will be operated on. This allows completely differently structured documents
to be contained within the same collection (however, in a well-designed Mongo database, multiple collections would be used).

@@

Often, the sensor readings received by the IRT team contains noisy data, defined by impossible values. For example, often
the kilometre tracking sensor can return distance values consisting of negative numbers. These are deemed to be bad
readings, and thus should not be included in any queries that are performed on the data. Hence, this data needs to be
excluded from the overall dataset in the MongoDB collection.

To exclude such data, it is possible to use MongoDB's provided \texttt{update()} function, which allows the selection
of noisy data based on a specified predicate. After data has been selected, \texttt{update()} allows the use of the
\texttt{\$unset} operator to remove specified fields from documents. This allows the other sensor readings in the same
document to persist in the database, while only removing those readings specified as noisy by the predicate.

This is performed through the following code:

\begin{lstlisting}[caption=MongoDB \texttt{update()} functions to remove data deemed to be noisy from existing documents.]
// Unset the bad km values (those with values less than 0)
db.train.update(
  { km: {$lt: 0} },       // predicate
  { $unset: {km: 1} },    // operation to perform
  { multi: true }         // option to specify operation on _all_ matching documents
);

// Unset all bad speed values (those with values less than 0)
db.train.update(
  { speed: {$lt: 0} },
  { $unset: {speed: 1} },
  { multi: true }
);

// Unset all bad LoadEmpty values (those not in valid enum values)
db.train.update(
  { LoadEmpty: {$not: {$in: ['unknown', 'empty', 'loaded']} } },
  { $unset: {LoadEmpty: 1} },
  { multi: true }
);

// Unset all bad Track values (those not in range of [0,213841])
db.train.update(
  { $or: [ {Track: {$lt: 0}}, {Track: {$gt: 213841}} ] },
  { $unset: {Track: 1} },
  { multi: true }
);
\end{lstlisting}

@@

\subsection{Ad-hoc adding of additional sensors} % (fold)
\label{sub:adding_additional_sensors}

In the case of the IRT project, it is very possible that additional sensors could be added at any future stage. For a
relational database, this would require modification to the table schemas, and possible restructuring operations
to be performed on existing tables to set values in the columns for the additional sensors. In MongoDB, as documents are
schema-less, new datasets containing new sensor's data can be serialised as BSON and inserted into existing collections
at will, without requiring any modifications to the underlying database.

As stated previously, differently structured documents can co-exist in the same collection. Hence, adding datasets
with additional sensors will simply result in documents being inserted that contain a different number of fields than
those documents that were previously stored in that collection.

For testing, an additional dataset has been created, based off the original dataset, with the addition of two extra
sensor fields: \textbf{Temp} and \textbf{Weight}. Documents with the addition of these new sensors simply have two
additional key-value pairs that are not present in those documents consisting of data from the original dataset. Any
queries performed that depend on either the \textbf{Temp} or \textbf{Weight} keys will simply only use those documents
where such values exist for these keys, thus ignoring documents from the original dataset, as touched on in~\sectref{sub:handling_missing_values}.

An example document containing these new fields is shown as follows:

\begin{lstlisting}[caption=Example of a document with extra sensor fields that do not exist in original dataset.]
{
  "_id" : ObjectId("556291145946aa793572b39a"),
  "Time" : 1.09447e+09,
  "Speed" : 4.4639997480000000,
  "LoadEmpty" : "unknown",
  "km" : 5.0096998220000000,
  "Track" : 2,
  "CouplerForce" : 23.1613006199999987,
  "Weight" : 2380,   // new pair
  "Temp" : 77,       // new pair
  "Loc" : {
    "Lat" : -20.6665759599999994,
    "Lon" : 117.1280811000000028
  },
  "Bounce" : {
    "Front" : -0.0303000000000000,
    "Rear" : -0.0029600000000000
  },
  "Rock" : {
    "Front" : 0.0383000000000000,
    "Rear" : 0.0062500000000000
  },
  "SND" : {
    "SND1" : -0.0111000000000000,
    "SND2" : 0.0001660000000000,
    "SND3" : undefined,
    "SND4" : -0.0494000000000000
  },
  "Acc" : {
    "Left" : 0.0000000000000000,
    "Right" : 0.0000000000000000
  }
}
\end{lstlisting}

% subsection adding_additional_sensors (end)

@@

As most of what has been shown in~\sectref{sec:implementation}, properly setting up the MongoDB database for use with the IRT
project is not a hard task and only requires relatively trivial MongoDB operations to be performed.
